---
title: "Assignment_04"
author: "Priyasha Chawla and Beryl Nana Ama Akuffo-Kwapong"
date: "17/02/2022"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)

```

```{r include = FALSE}
def.chunk.hook  <- knitr::knit_hooks$get("chunk")
knitr::knit_hooks$set(chunk = function(x, options) {
  x <- def.chunk.hook(x, options)
  ifelse(options$size != "normalsize", paste0("\n \\", options$size,"\n\n", x, "\n\n \\normalsize"), x)
})

```

\begin{center}
\Huge{PPOL 670 | Assignment 04}

\Huge{Data Visualization, Markdown, and Git}
\end{center}

\vspace{0.1in}

```{r}
library(tidyverse)
library(ggplot2)
library(reshape2)
library(forcats)
library(dplyr)
library(plotly)
library(viridis)
library(hrbrthemes)
library(scales)

library(httr)
set_config(config(ssl_verifypeer = 0L))
library(readxl)

project <- read_excel("AIU All Women Dataset.xlsx")

```


```{r}
#Graph 1 - Total maternal deaths per 100,000 live births
graph1 <- project %>%
  group_by(region) %>%
  summarise(weighted = weighted.mean(rate_matdeaths))

#Creates a line plot which shows the rate of maternal deaths in each continent(region)
graph1 %>%  
  ggplot(aes(x=region, y=weighted, group = 1)) +
  geom_point(size=3, colour = "red") +
  geom_line(colour = "blue") +
  labs(title = "Maternal deaths per 100,000 live births by region",
       caption = "Source: The Guttmacher Institute",
       x = "Region", y = "Maternal deaths per 100,000 live births") +
  theme_minimal()
```


Graph 1: Maternal deaths per 100,000 live births by region

It gives a glimpse of the maternal deaths per 100,000 live births by each region. Africa has the highest maternal death rate at 4 deaths per 100,000 live births. The regions with second highest maternal death rate are Latin America and the Caribbean and Asia.  Europe has the lowest maternal death rate. This could be a direct function of the economic development in these regions and, at the same time, leads to severe economic burden on households and negative impact on the Gross Domestic Product.  Research suggests that maternal mortality could lead to fall in GDP by $0.36 per year (Kirgia et all, 2006). This dual cause and effect relationship could worsen any economic advancement effects by developing countries.
[Source: The Guttmacher Institute](osf.io/m85s9)


